---
title: "Matching and Propensity Scores"
author: "Travis Loux"
date: '`r format(Sys.Date(), format="%b %d, %Y")`'
output:
  beamer_presentation:
    incremental: no
    fig_width: 4
    fig_height: 3
  ioslides_presentation:
    incremental: yes
    widescreen: yes
---

## Matching and propensity scores

- Why match?
- Matching methods
- Propensity scores
- Matching in `R`
- Diagnostics
- Analysis of matched data


## Why match?

- Matching pairs units in the treatment group to units in the control group
    - Numerous ways to do this

- Subjects not similar enough to be matched are discarded


## Why match?

- Matching can be thought of as **pre-processing** your data
- Yields treatment and control groups much more similar
    - Exchangeable


## Why match?

- After matching, treatment should be approximately independent of covariates
    - Similar to randomized studies, but only for observed variables

- Estimates of treatment effect are less model-dependent


## Why match?

[Shiny example]


## Matching methods

- Matching methods have two main components
    - **Distance**: computing how far subjects are away from each other
    - **Matching algorithm**: how subjects are paired together


## Matching methods

**Euclidean distance** measures the distance between two subjects in the measurement units

```{r, echo=FALSE, fig.align='center'}
x = rnorm(n=50, mean=5, sd=2)
y = rnorm(n=50, mean=5, sd=1)

par(mar=c(5,4,1,1))
plot(x, y, xlim=c(0,13), ylim=c(0,10), frame=FALSE)
points(x=c(2,6), y=c(3,6), col='blue', pch=19, cex=1.2)
lines(x=c(2,6), y=c(3,6), col='blue', lwd=2)
d_eu = sqrt((2-6)^2 + (3-6)^2)
text(x=8, y=8, labels=paste0("dist = ",round(d_eu,2)), cex=1.5)
```


## Matching methods

**Mahalanobis distance** measured the distance between two subjects in standardized units

- Accounts for different variances and correlations

```{r, echo=FALSE, fig.align='center'}
par(mar=c(5,4,1,1))
plot(x, y, xlim=c(0,13), ylim=c(0,10), frame=FALSE)
points(x=c(2,6), y=c(3,6), col='blue', pch=19, cex=1.2)
lines(x=c(2,6), y=c(3,6), col='blue', lwd=2)
d_mal = sqrt(((2-6)/2)^2 + (3-6)^2)
text(x=8, y=8, labels=paste0("dist = ",round(d_mal,2)), cex=1.5)
```


## Matching methods

- **One-to-one**: one treatment to one control
- **One-to-m**: one treatment to $m$ controls
- **One-to-many**: one treatment to a varying number of controls
- **Many-to-many**: varying numbers of treatments and controls
    - sometimes called **full matching**


## Matching methods

- **Exact**: only matches subjects with perfect concordance
- **Nearest neighbor**: matches treatment to "closest" control
    - **Greedy**: finds the best match at the given point
    - May lead to unmatchable subjects later
- **Optimal**: finds the set of matches that minimizes the total distance between all matched pairs


## Matching methods

- **Calipers** or **matching windows**: a limit on how far apart subjects can be and still be considered "matchable"
    - Exact matching has a caliper of 0
    - Difficult to determine in high-dimensional space

- **Bias-variance trade off**


## Propensity scores

The **propensity score** is one tool to get around the high-dimensional problem in matching

- Defined as probability of treatment conditional on a set of covariates
$$e(x) = P(T = 1 \mid X=x)$$

- The PS is a **balancing score**: any variable controlled for in $e(x)$ is balanced between treatment and control group
    - The coarsest balancing score


## Propensity scores

Benefits of using propensity score

- Dimension reduction: reduces a complex set of variables to a single number
    - Between 0 and 1
- Accounts for differences in scaling across variables
- Gives a quantitative difference between levels of categorical variables


## Propensity scores

Limitations of using propensity score

- Dimension reduction loses information
- Reliant on model assumptions
    - Not as much as outcome regression
- Does not balance excluded variables
    - Worse balance?


## Propensity scores

- The range of the propensity score is 0 to 1
    - Bounded differences
    - Especially problematic at the extremes
- Can use **linear propensity score**
    - Often by taking logit
    - $log\left( \frac{p}{1-p}\right)$


## Propensity scores

```{r, echo=FALSE, fig.align='center'}
library(boot)
curve(logit(x), main='Linear Propensity score', frame=FALSE)
```


## Matching in `R`

- The most common tool for matching in `R` is the **MatchIt** package
    - `install.packages('MatchIt')` - only once
    - `library(MatchIt)` - every R session

- For detailed help, <http://gking.harvard.edu/matchit>


## Matching in `R`

- Question: Are the sentences for crack cocaine longer than those for powder cocaine?
- Data: Monitoring of Federal Criminal Sentences, 2010
    - United States Sentencing Commission

- Build a DAG for this scenario


## Matching in `R`

```{r}
library(MatchIt)
load('drugs.RData')
nrow(drugs)
table(drugs$drug)
```


## Matching in `R`

To match data, use **matchit()** function

- Needs to have a binary treatment indicator
- Needs complete data (no missing values)


## Matching in `R`

```{r}
drugs$crack = as.numeric(drugs$drug == 'Crack')

drugs_complete = na.omit(drugs)
nrow(drugs_complete)

matching = matchit(crack ~ age + circuit + sex + 
                     citizen + educ + race, 
                   data=drugs_complete, 
                   method='nearest', 
                   distance='logit', 
                   caliper=0.10)
```


## Diagnostics

Important diagnostics to check:

- Number of treated left unmatched
    - More important than controls
- Distributions of covariates before and after matching
    - Means
    - Distribution of propensity scores
    

## Diagnostics

- For numerical diagnostics, use `summary(matching)`
- For graphical diagnostics, use `plot(matching)`
    - `plot(matching)` gives qq plots for all covariates
    - `plot(matching, type='jitter')` gives dot plots of PS
    - `plot(matching, type='hist')` gives histograms of PS
    - `plot(summary(matching,standardize=TRUE))`


## Analysis of matched data

To extract the matched data set, use **match.data()** function:

```{r}
match_drugs = match.data(matching, distance='ps')

nrow(match_drugs)
table(match_drugs$drug)
```


## Analysis of matched data

The new data set, `match_drugs`, can be analyzed as a standard data set

- $t$ tests or odds ratios
- Regression methods
- Non-parametric estimates


## Analysis of matched data

```{r}
sentence_model = lm(sentence ~ age + circuit + sex + 
                      citizen + educ + race + drug, 
                    data=match_drugs)

sent_sum = summary(sentence_model)
round(sent_sum$coefficients['drugCrack', ], 2)
```


## Analysis of matched data

```{r}
alt_model = glm(alternative ~ age + circuit + sex + 
                  citizen + educ + race + drug, 
                family=binomial('logit'), data=match_drugs)

alt_sum = summary(alt_model)
round(alt_sum$coefficients['drugCrack', ], 2)
exp(alt_sum$coefficients['drugCrack', 'Estimate'])
```


## Analysis of matched data

In matching, we usually start with the treatment group and find a control for each treated subject

- Covariate distributions look like the treatment group
- Will lead to an estimate of ATT
- Good estimate to the extent treated subjects were dropped
    - Will change covariate distribution of remaining subjects
    - Dropping too many treated subjects results in the population being ill-defined
