---
title: "Inverse Probability Weighting"
author: "Travis Loux"
date: '`r format(Sys.Date(), format="%b %d, %Y")`'
output:
  ioslides_presentation:
    incremental: yes
    widescreen: yes
  beamer_presentation:
    fig_height: 3
    fig_width: 4
    incremental: no
---

## Inverse probability weighting

- Propensity scores
- IPW
- Weighting diagnostics
- Standard errors and confidence intervals


# Propensity scores


## Propensity scores

- Probability of treatment given observed covariates
- $$e(x) = P(T=1 \mid X=1)$$

- Treated and control with same propensity score come from balanced populations
    - Distribution of covariates same across treatment groups
    - No confounding by $X$


## Propensity scores

```{r, echo=FALSE}
tr = rep(0:1, each=100)
x1 = rnorm(n=200, mean=tr/2)
x2 = rnorm(n=200, mean=-tr/4)
x3 = rnorm(n=200, mean=tr)

pscr = glm(tr ~ x1 + x2 + x3, family=binomial('logit'))$fitted

tr_jit = jitter(tr)
plot(pscr, tr_jit, xlim=0:1, frame=FALSE, 
     xlab='Propensity score', 
     yaxt='n', ylab='Treatment group')
axis(side=2, at=0:1, labels=c('Control','Treatment'), lwd=0)
```


## Propensity scores

```{r}
load('drugs.RData')
drugs$crack = as.numeric(drugs$drug == 'Crack')
drugs_complete = na.omit(drugs)

ps_model = glm(crack ~ age + circuit + sex + 
                 citizen + educ + race, 
               family=binomial('logit'), 
               data=drugs_complete)
ps = ps_model$fitted
```


# IPW


## IPW

- Besides matching, PS can be used for **inverse probability weighting** (**IPW**)
    - Also callsed inverse probability of treatment weighting (IPTW)

- Individuals can be weighted to match any population
    - Can estimate ATE or ATT depending on weights
    - Or the average effect in any other group


## IPW

Consider units with $e(x) = 0.20$

- Every unit at this PS level has a 0.20 probability of being treated
    - 1 in 5 are treated

- Every unit at this PS level has a 0.80 probability of being untreated
    - 4 in 5 are untreated

- Treated units are outnumbered 4 to 1


## IPW (ATE)

**ATE** measures the effect in the entire population (represented by sample)

- $ATE = E(Y^{(1)}) - E(Y^{(0)})$

- Each treatment group is weighted to the entire sample at that PS level


## IPW (ATE)

To estimate $E(Y^{(1)})$:

- Give units weight of $w^{(1)} = \frac{t}{e(x)}$
    - In treatment group:
    - $\frac{1}{e(x)} = \frac{1}{0.2} = 5$
    - One treated unit counts for 5 sample units

- $\frac{\sum_{i} w^{(1)}_i Y_i}{\sum_{i} w^{(1)}_i}$ estimates $E(Y^{(1)})$


## IPW (ATE)

To estimate $E(Y^{(0)})$:

- Give units weight of $w^{(0)} = \frac{1 - t}{1 - e(x)}$
    - In control group:
    - $\frac{1 - 0}{1 - 0.2} = \frac{1}{0.8} = 5/4$
    - Four control units count for a total of 5 sample units

- $\frac{\sum_{i} w^{(0)}_i Y_i}{\sum_{i} w^{(0)}_i}$ estimates $E(Y^{(0)})$


## IPW (ATE)

- Estimate the ATE by comparing weighted averages:

- $$ATE = E(Y^{(1)}) - E(Y^{(0)})$$

- $$\widehat{ATE}_{IPW} = 
\frac{\sum_{i} \frac{t_i}{e(x_i)} \times Y_i}{\sum_{i} \frac{t}{e(x_i)}} - 
\frac{\sum_{i} \frac{(1 - t_i)}{1 - e(x_i)} \times Y_i}{\sum_{i} \frac{1 - t}{1 - e(x_i)}}$$


## IPW (ATE)

```{r}
w_1_ate = drugs_complete$crack / ps
w_0_ate = (1 - drugs_complete$crack) / (1 - ps)
```


## IPW (ATE)

Compute weighted averages:

```{r}
(y1_ate = weighted.mean(x=drugs_complete$sentence, 
                    w=w_1_ate))
(y0_ate = weighted.mean(x=drugs_complete$sentence, 
                    w=w_0_ate))
(ate = y1_ate - y0_ate)
```


## IPW (ATT)

**ATT** measures the effect of treatment on the treated population

- $ATT = E(Y^{(1)} \mid T=1) - E(Y^{(0)} \mid T=1)$

- Each treatment group is weighted to the treated sample at that PS level


## IPW (ATT)

To estimate $E(Y^{(1)} \mid T=1)$:

- Give units weight of $t_i$
    - Treatment group weights 1

- $\frac{\sum_{i} t_i \times Y_i}{\sum_{i} t_i}$ estimates $E(Y^{(1)} \mid T=1)$
    - Sample mean of treated


## IPW (ATT)

To estimate $E(Y^{(0)} \mid T=1)$:

- Give units weight of $w^{(0)} = \frac{(1 - t_i) \times e(x)}{1 - e(x)}$
    - In control group:
    - $\frac{e(x)}{1 - e(x)} = \frac{0.2}{0.8} = 1/4$
    - Four control units count for a total of 1 treated unit

- $\frac{\sum_{i} w^{(0)}_i Y_i}{\sum_{i} w^{(0)}_i}$ estimates $E(Y^{(0)} \mid T=1)$
    - Counterfactual mean


## IPW (ATT)

- Estimate the ATT by comparing weighted averages:

- $$ATT = E(Y^{(1)} \mid T=1) - E(Y^{(0)} \mid T=1)$$

- $$\widehat{ATT}_{IPW} = 
\frac{\sum_{i} t_i \times Y_i}{\sum_{i} t_i} - 
\frac{\sum_{i} w^{(0)}_i Y_i}{\sum_{i} w^{(0)}_i}$$


## IPW (ATT)

```{r}
w_1_att = drugs_complete$crack
w_0_att = (1 - drugs_complete$crack) * ps / (1 - ps)
```


## IPW (ATT)

Compute weighted averages:

```{r}
(y1_att = weighted.mean(x=drugs_complete$sentence, 
                    w=w_1_att))
(y0_att = weighted.mean(x=drugs_complete$sentence, 
                    w=w_0_att))
(att = y1_att - y0_att)
```


## IPW

Summary:


Parameter      $w^{(1)}$             $w^{(0)}$
-----------  --------------------  ---------------------------
ATE            $\frac{t}{e(x)}$      $\frac{1 - t}{1 - e(x)}$
ATT            $t$                   $\frac{(1 - t) \times e(x)}{1 - e(x)}$


# Weighting diagnostics


## Weighting diagnostics

- Weights works well as long as they are not too large

- Large weights:
    - Increase standard errors
    - Indicate positivity problems

- The most common diagnostics
    - Check treatment group weight total
    - Check extreme weights


## Weighting diagnostics

- Weights should total to the size of the relevant sample
    - ATE: total sample size
    - ATT: treatment group sample size


## Weighting diagnostics

Using ATE weights:

```{r}
nrow(drugs_complete)

sum(w_1_ate)
sum(w_1_ate) / nrow(drugs_complete)
```


## Weighting diagnostics

Using ATE weights:

```{r}
nrow(drugs_complete)

sum(w_0_ate)
sum(w_0_ate) / nrow(drugs_complete)
```


## Weighting diagnostics

Using ATT weights:

```{r}
sum(drugs_complete$crack==1)

sum(w_0_att)
sum(w_0_att) / sum(drugs_complete$crack==1)
```

- Why not check `w_1_att`?


## Weighting diagnostics

- "Uncommon" subjects are given larger weights
    - Treated units with $e(x) \approx 0$
    - Control units with $e(x) \approx 1$

- Example: if a treated unit has $e(x) = 0.01$, then $w^{(1)} = \frac{1}{0.01} = 100$
    - A single treated unit accounts for 100 sample units
    - A small change in $e(x)$ yield a large change in weight
        - $e(x) = 0.02 \rightarrow w^{(1)} = 50$


## Weighting diagnostics

```{r}
summary(w_1_ate[drugs_complete$crack==1])
sort(w_1_ate, decreasing=TRUE)[1:10]
```


## Weighting diagnostics

```{r}
summary(w_0_ate[drugs_complete$crack==0])
sort(w_0_ate, decreasing=TRUE)[1:10]
```


## Weighting diagnostics

- A common approach for handling extreme weights is **weight truncation**
    - Large weights are capped to some pre-determined value
    - Usually 10-20

- Another bias-variance tradeoff
    - PS weighting is unbiased (given assumptions)
    - Large weights lead to large variance
    - Artificially truncating weights will reduce variance, but add bias


## Weighting diagnostics

To cap weights at 15:

```{r}
w_1_trunc = pmin(w_1_ate, 15)
w_0_trunc = pmin(w_0_ate, 15)

nrow(drugs_complete)

sum(w_1_trunc)
sum(w_0_trunc)
```


## Weighting diagnostics

```{r}
(y1_trunc = weighted.mean(x=drugs_complete$sentence, 
                          w=w_1_trunc))
(y0_trunc = weighted.mean(x=drugs_complete$sentence, 
                          w=w_0_trunc))
(ate_trunc = y1_trunc - y0_trunc)
```


## Standard errors and confidence intervals

- SEs and CIs are not as straightforward
- Weights are data-dependent
    - Random
    - Correlated with $Y$

- Two choices:
    - Use estimates from survey weighting (sandwich estimator)
        - Tends to be over-conservative
    - Bootstrap
        - Computationally difficult


## Standard errors and confidence intervals

- To use survey weighting methods, need **survey** package
    - `install.packages('survey')`
    - `load(survey)`

- Needs to create a survey object that includes weights


## Standard errors and confidence intervals

```{r, message=FALSE, warning=FALSE}
library(survey)

w_ate = w_1_ate + w_0_ate

drugs_survey = svydesign(ids=~0, 
                         data=drugs_complete, 
                         weights=w_ate)
```


## Standard errors and confidence intervals

To get weighted estimate, run a regression of outcome on treatment

```{r}
ate_model = svyglm(sentence ~ crack, 
                   design=drugs_survey)
summary(ate_model)$coefficients

c(17.35 - 1.96*4.45, 17.35 + 1.96*4.45)
```


## Standard errors and confidence intervals

- For binary outcome data, include **family=binomial('logit')** for logistic regression
- To get CI for the odds ratio:
    - Estimate $\pm 1.96\times$ standard error
    - Exponentiate (**exp()**)


## Standard errors and confidence intervals

- To bootstrap:
    1. Write a function that will resample data and compute weighted estimate
    2. Apply function to data large number of times (10,000)
    3. Middle 95% of estimates make up CI


## Standard errors and confidence intervals

(See RMarkdown file for bootstrap code)

```{r, echo=FALSE}
ate_boot = function(x){
  
  # sample data
  boot_index = sample.int(nrow(x), size=nrow(x), 
                          replace=TRUE)
  boot_data = x[boot_index, ]
  
  # create ps
  boot_ps_model = glm(crack ~ age + circuit + sex + 
                        citizen + educ + race, 
                      family=binomial('logit'), 
                      data=boot_data)
  boot_ps = boot_ps_model$fitted
  
  # create weights
  boot_wt_1 = boot_data$crack / boot_ps
  boot_wt_0 = (1 - boot_data$crack) / (1 - boot_ps)
  
  # estimate ATE
  boot_y1 = weighted.mean(x=boot_data$sentence, w=boot_wt_1)
  boot_y0 = weighted.mean(x=boot_data$sentence, w=boot_wt_0)
  boot_y1 - boot_y0
}
```


## Standard errors and confidence intervals

```{r}
boot_ates = replicate(n=100, 
                      expr=ate_boot(x=drugs_complete))
hist(boot_ates, main='Bootstrap histogram', 
     xlab='ATE estimate')
```


## Standard errors and confidence intervals

```{r}
quantile(boot_ates, probs=c(0.025, 0.975))
```
