---
title: "Correlated Counfounder and Propensity Score Matching"
author: "Miao Cai"
date: '`r Sys.Date()`'
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Corrections in this 20171202 version

## 1 Keep the directions of Cohen's d consistent across all comparisons^[This is correctified by explicitly assigning the levels of the factors `levels = c(0, 1)` in `cohen.d(dat$y,factor(dat$tr, levels = c(0, 1)))$estimate`.]

## 2 Cohen's d for all 9 covariates, for 3 covariates and 3 PCs

## 3 Adding the 7 covariates and 5 covariates part

## 3 Display the loading factors

## 4 Adding the Sufficient Reduction Method

\newpage

# Creating simulation data
Random variables $X_1$ - $X_3$ have the correlation coefficient of 0.3; random variables $X_4$ - $X_6$ have the correlation coefficient of 0.5; random variables $X_7$ - $X_9$ have the correlation coefficient of 0.8. The true population parameters for $X_1, X_4, X_7$ is 2, parameters for $X_2, X_5, X_8$ is 3, parameters for $X_3, X_6, X_9$ is 1.
```{r ,cache=FALSE}
library(MASS)
library(Matrix)
library(GMCM)
library(MatchIt)
set.seed(666)

# correlations
r1 = 0.3
r2 = 0.5
r3 = 0.8

# block diagnoal correlation matrix
m1 = matrix(r1, nrow=3, ncol=3)
diag(m1) = 1
m2 = matrix(r2, nrow=3, ncol=3)
diag(m2) = 1
m3 = matrix(r3, nrow=3, ncol=3)
diag(m3) = 1

cmat = bdiag(m1, m2, m3)

# covariates
x = data.frame(mvrnorm(n=1000, mu=rep(0,9), Sigma=cmat))

# pt: the probability to draw the binary treatment
##REVISED: rowSums(x)-3.8 to reduce proportion treated
pt = GMCM:::inv.logit(rowSums(x)-3.8)
## REVISED: to confirm that mean(pt) is near 0.2
mean(pt) 
# mean(pt) is around 0.2 to make sure there are sufficient 
# number of comparison groups to choose from.

# tr: treatment
tr = rbinom(n = 1000, size = 1, prob = pt)

# y: outcome - POPULATION PARAMETER for treatment is 3
y = rnorm(n = 1000, 
          mean = tr * 3 + 3*x$X1 + 2*x$X2 + x$X3 + 3*x$X4 + 2*x$X5 + x$X6 + 3*x$X7 + 2*x$X8 + x$X9, 
          sd = 1)

# constructing the data.frame
dat <- data.frame(x, tr, y)
```
\newpage

# Part 1 Nine Covariates
This part firstly uses all 9 correlated covariates to match the treatment and comparison group^[Propensity score method is used to match the treatment group and the comparison group. I use the [**MatchIt** package](https://github.com/cran/MatchIt) to do propensity score matching]. Then I use linear regression to estimate the coefficients of $X_1 \sim X_9$, and Cohen's d is used to test the effect size.^[Cohen's d is calculated using the following formula:$$Cohen's \ d = \frac{\bar{X_1} - \bar{X_2}}{\sqrt{\frac{(n_1-1)s_1^2 + (n_2 - 1)s_2^2}{n_1 + n_2 - 2}}}$$ ,with the _cohen.d()_ function in the [**effsize** package](https://github.com/cran/effsize). When paired is set, the effect size is computed using the approach suggested in (Gibbons et al. 1993) _Gibbons, R. D., Hedeker, D. R., & Davis, J. M. (1993). Estimation of effect size from a series of experiments involving paired comparisons. Journal of Educational Statistics, 18, 271-279._]

## Section 1.1 Nine covariates without matching
## 1.1.1 `y ~ tr` on unmatched data
```{r}
library(effsize)
lm1.1.1 <- lm(y ~ tr, data = dat)

# summary the output
knitr::kable(
  summary(lm1.1.1)$coefficients, 
  caption = 'Linear regression between y and treatment on unmatched data', 
  digits = 3
)

# get the Cohen's d for this model
cohen.d(dat$y,factor(dat$tr, levels = c(0, 1)))$estimate
```

## 1.1.2 `y ~ tr + 9` covariates on unmatched data
```{r}
lm1.1.2 <- lm(y ~ tr + X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9,
              data = dat)

# summary the output
knitr::kable(
  summary(lm1.1.2)$coefficients, 
  caption = 'Linear regression between y and treatment, 9 covariates on unmatched data', 
  digits = 3
)
```

## 1.1.3 Cohen's d for each covariate by tr
```{r}
cohen.d(dat$X1,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X2,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X3,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X4,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X5,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X6,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X7,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X8,factor(dat$tr, levels = c(0, 1)))$estimate
cohen.d(dat$X9,factor(dat$tr, levels = c(0, 1)))$estimate
```


##Section 1.2 Nine covariates with matching
## 1.2.1 `y ~ tr` on matched data
```{r}
#1 match the treatment and comparison groups - 1 to 1 match
matcheddata1 <- match.data(
  matchit(tr ~ X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9, 
          data = dat, 
          method = "nearest", 
          ratio = 1))

#2 linear regression - y and treatment on matched data
lm1.2.1 <- lm(y ~ tr, data = matcheddata1)
knitr::kable(
  summary(lm1.2.1)$coefficients, 
  caption = 'Linear regression between y and treatment on matched data', 
  digits = 3
)

cohen.d(matcheddata1$y, factor(matcheddata1$tr, levels = c(0, 1)))$estimate
```

## 1.2.2 `y ~ tr + 9` covariates on matched data
```{r}
#3.2 linear regression - y, treatment and covariates
lm1.2.2 <- lm(y ~ tr + X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9, 
            data = matcheddata1)
knitr::kable(
  summary(lm1.2.2)$coefficients, 
  caption = 'Linear regression between y ,treatment and 9 covariates', 
  digits = 3
)

#4 effect size - Cohen's d
cohen.d(matcheddata1$y, factor(matcheddata1$tr, levels = c(0, 1)))$estimate
```

## 1.2.3 Cohen's d for each covariate by tr
```{r}
cohen.d(matcheddata1$X1,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X2,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X3,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X4,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X5,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X6,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X7,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X8,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata1$X9,factor(matcheddata1$tr, levels = c(0, 1)))$estimate
```

\newpage

# Part 2 Seven Covariates (X1 ~ X7)
This part compares the results regressing $X_1$ - $X_7$ on y and the regression after propensity score matching.

## Section 2.1 Seven covariates without matching
## 2.1.1 `y ~ tr + X1~X7` on unmatched data
```{r}
library(effsize)
lm2.1.1 <- lm(y ~ tr + X1 + X2 + X3 + X4 + X5 + X6 + X7,
              data = dat)

# summary the output
knitr::kable(
  summary(lm2.1.1)$coefficients, 
  caption = 'Linear regression between y and treatment, 7 covariates on unmatched data', 
  digits = 3
)

# get the Cohen's d for this model
cohen.d(dat$y,factor(dat$tr, levels = c(0, 1)))$estimate
```

## Section 2.2 Seven covariates with matching
## 2.2.1 `y ~ tr` on matched data
```{r}
#1 match the treatment and comparison groups - 1 to 1 match
matcheddata2 <- match.data(
  matchit(tr ~ X1 + X2 + X3 + X4 + X5 + X6 + X7, 
          data = dat, 
          method = "nearest", 
          ratio = 1))

#2 linear regression - y and treatment on matched data
lm2.2.1 <- lm(y ~ tr, data = matcheddata2)
knitr::kable(
  summary(lm2.2.1)$coefficients, 
  caption = 'Linear regression between y and treatment on matched data',   digits = 3
)

cohen.d(matcheddata2$y, factor(matcheddata2$tr, levels = c(0, 1)))$estimate
```

## 2.2.2 `y ~ tr + X1~X7` on matched data
```{r}
#2.2 linear regression - y, treatment and 7 covariates
lm2.2.2 <- lm(y ~ tr + X1 + X2 + X3 + X4 + X5 + X6 + X7, 
            data = matcheddata2)
knitr::kable(
  summary(lm2.2.2)$coefficients, 
  caption = 'Linear regression between y ,treatment and 7 covariates', 
  digits = 3
)

#4 effect size - Cohen's d
cohen.d(matcheddata2$y, factor(matcheddata2$tr, levels = c(0, 1)))$estimate
```

## 2.2.3 Cohen's d for each covariate by tr
```{r}
cohen.d(matcheddata2$X1,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X2,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X3,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X4,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X5,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X6,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X7,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X8,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata2$X9,factor(matcheddata2$tr, levels = c(0, 1)))$estimate
```
\newpage

# Part 3 Five Covariates (X1 ~ X4, X7)
This part compares the results regressing $X_1$ - $X_4$ and $X_7$ on y and the regression after propensity score matching.

## Section 3.1 Five covariates without matching
## 3.1.1 `y ~ tr + X1~X4 + X7` on unmatched data
```{r}
library(effsize)
lm3.1.1 <- lm(y ~ tr + X1 + X2 + X3 + X4 + X7,
              data = dat)

# summary the output
knitr::kable(
  summary(lm3.1.1)$coefficients, 
  caption = 'Linear regression between y and treatment, 5 covariates on unmatched data', 
  digits = 3
)

# get the Cohen's d for this model
cohen.d(dat$y,factor(dat$tr, levels = c(0, 1)))$estimate
```

## Section 3.2 Five covariates with matching
## 3.2.1 `y ~ tr` on matched data
```{r}
#1 match the treatment and comparison groups - 1 to 1 match
matcheddata3 <- match.data(
  matchit(tr ~ X1 + X2 + X3 + X4 + X7, 
          data = dat, 
          method = "nearest", 
          ratio = 1))

#2 linear regression - y and treatment on matched data
lm3.2.1 <- lm(y ~ tr, data = matcheddata3)
knitr::kable(
  summary(lm3.2.1)$coefficients, 
  caption = 'Linear regression between y and treatment on matched data',   digits = 3
)

cohen.d(matcheddata3$y, factor(matcheddata3$tr, levels = c(0, 1)))$estimate
```

## 3.2.2 `y ~ tr + X1~X7` on matched data
```{r}
# linear regression - y, treatment and 5 covariates
lm3.2.2 <- lm(y ~ tr + X1 + X2 + X3 + X4 + X7, 
            data = matcheddata3)
knitr::kable(
  summary(lm3.2.2)$coefficients, 
  caption = 'Linear regression between y ,treatment and 9 covariates', 
  digits = 3
)

# effect size - Cohen's d
cohen.d(matcheddata3$y, factor(matcheddata3$tr, levels = c(0, 1)))$estimate
```

## 3.2.3 Cohen's d for each covariate by tr
```{r}
cohen.d(matcheddata3$X1,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X2,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X3,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X4,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X5,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X6,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X7,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X8,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata3$X9,factor(matcheddata3$tr, levels = c(0, 1)))$estimate
```
\newpage

# Part 4 Three Covariates
This part firstly uses 3 correlated covariates to match the treatment and comparison group. Then propensity scores are used to match the treatment groups and comparison groups. Linear regression and Cohen's d are conducted after propensity score matching.

## Section 4.1 Three uncorrelated covariates on unmatched data
## 4.1.1 `y ~ tr + 3` uncorrelated covariates on unmatched data
```{r}
lm4.1.1 <- lm(y ~ tr + X1 + X4 + X7, data = dat)

# summary the output
knitr::kable(
  summary(lm4.1.1)$coefficients, 
  caption = 'Linear regression between y and treatment, 3 uncorrelated covariates on unmatched data', 
  digits = 3
)
```

## Section 4.2 Three uncorrelated covariates on matched data
## 4.2.1 `y ~ tr` on matched data
```{r}
#1 match the treatment and comparison groups - 1 to 1 match
matcheddata4 <- match.data(
  matchit(tr ~ X1 + X4 + X7, 
          data = dat, 
          method = "nearest", 
          ratio = 1))

#2 linear regression - y and treatment
lm4.2.1 <- lm(y ~ tr, data = matcheddata4)
knitr::kable(
  summary(lm4.2.1)$coefficients, 
  caption = 'Linear regression between y and treatment on matched data', 
  digits = 3
)

#3 Cohen's d
cohen.d(matcheddata4$y, factor(matcheddata4$tr, levels = c(0, 1)))$estimate
```

## 4.2.2 `y ~ tr + X1 + X4 + X7` on matched data
```{r}
#1 linear regression - y， treatment and covariates
lm4.2.2 <- lm(y ~ tr + X1 + X4 + X7, 
            data = matcheddata4)
knitr::kable(
  summary(lm4.2.2)$coefficients, 
  caption = 'Linear regression between y ,treatment and 3 covariates on matched data', 
  digits = 3
)

#2 Cohen's d
cohen.d(matcheddata4$y, factor(matcheddata4$tr, levels = c(0, 1)))$estimate
```

## 4.2.3 Cohen's d for each covariate by tr
```{r}
cohen.d(matcheddata4$X1,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X2,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X3,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X4,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X5,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X6,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X7,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X8,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata4$X9,factor(matcheddata4$tr, levels = c(0, 1)))$estimate
```
\newpage

# Part 5 Integrating 9 Covariates into 3 Principal Components
This part integrates the 9 covariates into 3 principal components using one principal component analysis.^[Prinpal component analysis is conducted using the base R function _prcomp()_] Then propensity scores are used to match the treatment groups and comparison groups using the 3 principal components. Linear regression and Cohen's d are conducted after propensity score matching.

## Section 5.1 Regression on unmatched data
## 5.1.1 Principle component analysis
```{r}
#1 principal component analysis
pca3 <- prcomp(dat[,paste("X", 1:9, sep = "")], scale = FALSE)
pca3data <- data.frame(
  dat$y, 
  dat$tr, 
  pca3$x[,1:3]
  )#extract the  three PCs, y and tr
names(pca3data) <- c("y", "tr", "PC1", "PC2", "PC3")

#2 The standard deviation of the principle components
knitr::kable(
  pca3$sdev, 
  caption = 'The standard deviation of the principle components', 
  digits = 3
)

#3 The matrix of variable loadings (columns are eigenvectors)
knitr::kable(
  pca3$rotation, 
  caption = 'The matrix of variable loadings (columns are eigenvectors)', 
  digits = 3
)

#4 The variable means
knitr::kable(
  pca3$center, 
  caption = 'The variable means', 
  digits = 3
)
```

## 5.1.2 `y ~ tr + 3PCs` on unmatched data
```{r}
#2 Linear regression on unmatched data
lm5.1.2 <- lm(y ~ tr + PC1 + PC2 + PC3, data = pca3data)

# summary the output
knitr::kable(
  summary(lm5.1.2)$coefficients, 
  caption = 'Linear regression between y and treatment and 3 PCs on unmatched data', 
  digits = 3
)

# get the Cohen's d for this model
cohen.d(pca3data$y, factor(pca3data$tr, levels = c(0, 1)))$estimate
```

## 5.1.3 Cohen's d for each covariate by tr
```{r}
cohen.d(pca3data$PC1,factor(pca3data$tr, levels = c(0, 1)))$estimate
cohen.d(pca3data$PC2,factor(pca3data$tr, levels = c(0, 1)))$estimate
cohen.d(pca3data$PC3,factor(pca3data$tr, levels = c(0, 1)))$estimate
```

## Section 5.2 Regression on matched data
## 5.2.1 `y ~ tr` on matched data
```{r}
#1 propensity score matching - one to one match
matcheddata5 <- match.data(
  matchit(tr ~ PC1 + PC2 + PC3, 
          data = pca3data, 
          method = "nearest", 
          ratio = 1))

#2 combine the PCs data (including PC1 - PC3) and original data (including X1 - X9)
matcheddata5 <- cbind(matcheddata5, dat[rownames(matcheddata5),-c(which(colnames(dat) == "y"), which(colnames(dat) == "tr"))])

#3 linear regression - y and treatment
lm5.2.1 <- lm(y ~ tr, data = matcheddata5)
knitr::kable(
  summary(lm5.2.1)$coefficients, 
  caption = 'Linear regression between y and treatment on matched data', 
  digits = 3
)

#4 Cohen's d
cohen.d(matcheddata5$y, factor(matcheddata5$tr, levels = c(0, 1)))$estimate
```

## 5.2.2 `y ~ tr + 3PC` on matched data
```{r}
#1 linear regression - y， treatment and covariates
lm5.2.2 <- lm(y ~ tr + PC1 + PC2 + PC3, 
            data = matcheddata5)
knitr::kable(
  summary(lm5.2.2)$coefficients, 
  caption = 'Linear regression between y ,treatment and 3 PCs on matched data', 
  digits = 3
)

#2 Cohen's d
cohen.d(matcheddata5$y, matcheddata5$tr)$estimate
```

## 5.2.3 Cohen's d for each covariate by tr
```{r}
#1 Cohen's d for each PCs by tr
cohen.d(matcheddata5$PC1,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$PC2,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$PC3,factor(matcheddata5$tr, levels = c(0, 1)))$estimate

#2 Cohen's d for each covariate by tr
cohen.d(matcheddata5$X1,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X2,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X3,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X4,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X5,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X6,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X7,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X8,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata5$X9,factor(matcheddata5$tr, levels = c(0, 1)))$estimate
```
\newpage

# Part 6 Separately Integrating 9 Covariates into 3 sets of Principal Components
This part separately integrates the 9 covariates into 3 sets principal components.^[Different from part 3, this part uses 3 principal component analyses and integrates $X_1 - X_3$ into $PC_1$, integrates $X_4 - X_6$ into $PC_2$, and integrates $X_7 - X_9$ into $PC_3$.] Then propensity scores are used to match the treatment groups and comparison groups using the 3 sets of principal components. Linear regression and Cohen's d are conducted after propensity score matching.

## Section 6.1 Regression on unmatched data
## 6.1.1 principal component analysis - 3 sets
```{r}
# principal component analysis - 3 sets
pca6.1 <- prcomp(dat[,paste("X", 1:3, sep = "")], scale = FALSE)
pca6.2 <- prcomp(dat[,paste("X", 4:6, sep = "")], scale = FALSE)
pca6.3 <- prcomp(dat[,paste("X", 7:9, sep = "")], scale = FALSE)
pca6data <- data.frame(
  dat$y, 
  dat$tr, 
  pca6.1$x[,1],
  pca6.2$x[,1],
  pca6.3$x[,1]
  )#extract the  three PCs, y and tr
names(pca6data) <- c("y", "tr", "PC1", "PC2", "PC3")

# Set 1: X1 ~ X3
## 1.1 The standard deviation of the principle components
knitr::kable(
  pca6.1$sdev, 
  caption = 'The standard deviation of the principle components set 1',
  digits = 3
)

## 1.2 The matrix of variable loadings (columns are eigenvectors)
knitr::kable(
  pca6.1$rotation, 
  caption = 'The matrix of variable loadings set1', 
  digits = 3
)

## 1.3 The variable means
knitr::kable(
  pca6.1$center, 
  caption = 'The variable means set 1', 
  digits = 3
)

# Set 2: X4 ~ X6
## 2.1 The standard deviation of the principle components
knitr::kable(
  pca6.2$sdev, 
  caption = 'The standard deviation of the principle components set 2',
  digits = 3
)

## 2.2 The matrix of variable loadings (columns are eigenvectors)
knitr::kable(
  pca6.2$rotation, 
  caption = 'The matrix of variable loadings set2', 
  digits = 3
)

## 2.3 The variable means
knitr::kable(
  pca6.2$center, 
  caption = 'The variable means set 2', 
  digits = 3
)

# Set 3: X7 ~ X9
## 3.1 The standard deviation of the principle components
knitr::kable(
  pca6.3$sdev, 
  caption = 'The standard deviation of the principle components set 3',
  digits = 3
)

## 3.2 The matrix of variable loadings (columns are eigenvectors)
knitr::kable(
  pca6.3$rotation, 
  caption = 'The matrix of variable loadings set 3', 
  digits = 3
)

## 3.3 The variable means
knitr::kable(
  pca6.3$center, 
  caption = 'The variable means set 3', 
  digits = 3
)
```

## 6.1.2 `y ~ tr + 3PCs` on unmatched data
```{r}
#2 Linear regression on unmatched data
lm6.1.2 <- lm(y ~ tr + PC1 + PC2 + PC3, data = pca6data)

# summary the output
knitr::kable(
  summary(lm6.1.2)$coefficients, 
  caption = 'Linear regression between y and treatment and 3 PCs on unmatched data', 
  digits = 3
)

# get the Cohen's d for this model
cohen.d(pca6data$y, factor(pca6data$tr, levels = c(0, 1)))$estimate
```

## 6.1.3 Cohen's d for each covariate by tr
```{r}
cohen.d(pca6data$PC1,factor(pca6data$tr, levels = c(0, 1)))$estimate
cohen.d(pca6data$PC2,factor(pca6data$tr, levels = c(0, 1)))$estimate
cohen.d(pca6data$PC3,factor(pca6data$tr, levels = c(0, 1)))$estimate
```

## Section 6.2 Regression on matched data
## 6.2.1 `y ~ tr` on matched data
```{r}
#1 propensity score matching
matcheddata6 <- match.data(
  matchit(tr ~ PC1 + PC2 + PC3, 
          data = pca6data, 
          method = "nearest", 
          ratio = 1))

#2 combine the PCs data (including PC1 - PC3) and original data (including X1 - X9)
matcheddata6 <- cbind(matcheddata6, dat[rownames(matcheddata6),-c(which(colnames(dat) == "y"), which(colnames(dat) == "tr"))])

#2 linear regression - y and treatment
lm6.2.1 <- lm(y ~ tr, data = matcheddata6)
knitr::kable(
  summary(lm6.2.1)$coefficients, 
  caption = 'Linear regression between y and treatment on matched data', 
  digits = 3
)

#3 Cohen's d
cohen.d(matcheddata6$y, factor(matcheddata6$tr, levels = c(0, 1)))$estimate
```

## 6.2.2 `y ~ tr + 3PC` on matched data
```{r}
#1 linear regression - y， treatment and covariates
lm6.2.2 <- lm(y ~ tr + PC1 + PC2 + PC3, 
            data = matcheddata6)
knitr::kable(
  summary(lm6.2.2)$coefficients, 
  caption = 'Linear regression between y ,treatment and 3 PCs on matched data', 
  digits = 3
)

#2 Cohen's d
cohen.d(matcheddata6$y, factor(matcheddata6$tr, levels = c(0, 1)))$estimate
```

## 6.2.3 Cohen's d for each covariate by tr
```{r}
#1 Cohen's d for each PCs by tr
cohen.d(matcheddata6$PC1,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$PC2,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$PC3,factor(matcheddata6$tr, levels = c(0, 1)))$estimate

#2 Cohen's d for each covariate by tr
cohen.d(matcheddata6$X1,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X2,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X3,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X4,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X5,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X6,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X7,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X8,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
cohen.d(matcheddata6$X9,factor(matcheddata6$tr, levels = c(0, 1)))$estimate
```

```{r digits, echo = FALSE}
options(digits=2)
```

# Summary Table of the above information
Statistics | No matching | Nine Covariates | Seven Covariates($X_1$ - $X_7$) | Five Covariates ($X_1$ - $X_4$, $X_7$) | Three Covariates ($X_1$, $X_4$, $X_7$) | 3 PCs(1 decomp) | 3 PCs (3 decomp)
----------- | ----------- | ----------- | ----------- | ----------- | ----------- | -----------
y ~ tr + covariates, tr coefficient (pre-match)|`r unname(lm1.1.2$coefficients[which(labels(lm1.1.2$coefficients)== "tr")])` | `r unname(lm1.1.2$coefficients[which(labels(lm1.1.2$coefficients)== "tr")])`| `r unname(lm2.1.1$coefficients[which(labels(lm2.1.1$coefficients)== "tr")])`| `r unname(lm3.1.1$coefficients[which(labels(lm3.1.1$coefficients)== "tr")])`| `r unname(lm4.1.1$coefficients[which(labels(lm4.1.1$coefficients)== "tr")])` | `r unname(lm5.1.2$coefficients[which(labels(lm5.1.2$coefficients)== "tr")])` | `r unname(lm6.1.2$coefficients[which(labels(lm6.1.2$coefficients)== "tr")])`
cohen's d, x1 ~ tr|`r unname(cohen.d(dat$X1,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X1,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X1,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X1,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X1,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X1,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X1,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x2 ~ tr|`r unname(cohen.d(dat$X2,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X2,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X2,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X2,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X2,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X2,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X2,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x3 ~ tr|`r unname(cohen.d(dat$X3,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X3,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X3,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X3,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X3,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X3,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X3,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x4 ~ tr|`r unname(cohen.d(dat$X4,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X4,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X4,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X4,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X4,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X4,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X4,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x5 ~ tr|`r unname(cohen.d(dat$X5,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X5,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X5,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X5,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X5,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X5,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X5,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x6 ~ tr|`r unname(cohen.d(dat$X6,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X6,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X6,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X6,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X6,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X6,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X6,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x7 ~ tr|`r unname(cohen.d(dat$X7,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X7,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X7,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X7,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X7,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X7,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X7,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x8 ~ tr|`r unname(cohen.d(dat$X8,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X8,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X8,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X8,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X8,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X8,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X8,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
cohen's d, x9 ~ tr|`r unname(cohen.d(dat$X9,factor(dat$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata1$X9,factor(matcheddata1$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata2$X9,factor(matcheddata2$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata3$X9,factor(matcheddata3$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata4$X9,factor(matcheddata4$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata5$X9,factor(matcheddata5$tr, levels = c(0, 1)))$estimate)`|`r unname(cohen.d(matcheddata6$X9,factor(matcheddata6$tr, levels = c(0, 1)))$estimate)`
y ~ tr, tr coef|`r unname(lm1.1.1$coefficients[which(labels(lm1.1.1$coefficients)== "tr")])`|`r unname(lm1.2.1$coefficients[which(labels(lm1.2.1$coefficients)== "tr")])`|`r unname(lm2.2.1$coefficients[which(labels(lm2.2.1$coefficients)== "tr")])`|`r unname(lm3.2.1$coefficients[which(labels(lm3.2.1$coefficients)== "tr")])`|`r unname(lm4.2.1$coefficients[which(labels(lm4.2.1$coefficients)== "tr")])`|`r unname(lm5.2.1$coefficients[which(labels(lm5.2.1$coefficients)== "tr")])`|`r unname(lm6.2.1$coefficients[which(labels(lm6.2.1$coefficients)== "tr")])`
y ~ tr + covariates, tr coef|`r unname(lm1.1.2$coefficients[which(labels(lm1.1.2$coefficients)== "tr")])` | `r unname(lm1.2.2$coefficients[which(labels(lm1.2.2$coefficients)== "tr")])`| `r unname(lm2.2.2$coefficients[which(labels(lm2.2.2$coefficients)== "tr")])`|`r unname(lm3.2.2$coefficients[which(labels(lm3.2.2$coefficients)== "tr")])`|`r unname(lm4.2.2$coefficients[which(labels(lm4.2.2$coefficients)== "tr")])`|`r unname(lm5.2.2$coefficients[which(labels(lm5.2.2$coefficients)== "tr")])`|`r unname(lm6.2.2$coefficients[which(labels(lm6.2.2$coefficients)== "tr")])`

































